---
title: "RNA_hostplant_DESeq2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r get_data}
#source("https://bioconductor.org/biocLite.R")
#biocLite("DESeq2")
#biocLite("sva")
#biocLite("cowplot")
library("DESeq2")
library(sva)
library(ggplot2)

sessionInfo()

#Clear all states
#rm(list=ls(all=TRUE))
#dev.off()

#This is a DEseq analysis of all experimental groups together, including PCA-analysis and visualisation.

#Load gene count matrix and labels ####

#setwd("")
countData <- as.matrix(read.csv("../outputs/2c_Quantification_StringTie_STEP2/gene_count_matrix.csv", row.names="gene_id"))
head(countData)
nrow(countData)

#setwd("/home/luisleal/MYPROJ/3_DosageCompensation_LS/Scripts_NEW")
colData <- read.csv("/proj/uppstore2017185/b2014034_nobackup/Karin/RNAseq_hostplant/working/sample_info_DEseq_corr2.txt", sep="\t", row.names=1, header=F)
colnames(colData) <- c("Hostplant", "Population", "Stage", "Sex", "Family", "Bodypart")
colData$Stage <- as.factor(colData$Stage)
colData
str(colData)

#Note: The PHENO_DATA file contains information on each sample, e.g., sex or population.
#The exact way to import this depends on the format of the file.




##### output folders

# standard filtering
#OUT_FOLD_STDF <- "/crex1/proj/uppstore2017185/b2014034_nobackup/Veronika/RNA_HostPlantProject/results/4_DESeq2/03_Interaction_effect/"


#From deseq2 manual:
#...for some datasets, exploratory data analysis (EDA) plots could reveal that one
#or more groups has much higher within-group variability than the others. A simulated
#example of such a set of samples is shown below. This is case where, by comparing 
#groups A and B separately - subsetting a DESeqDataSet to only samples from those two 
#groups and then running DESeq on this subset - will be more sensitive than a model 
#including all samples together. It should be noted that such an extreme range of 
#within-group variability is not common, although it could arise if certain treatments
#produce an extreme reaction (e.g. cell death). Again, this can be easily detected
#from the EDA plots such as PCA described in this vignette.





#Pre-filtering I: Only keep genes with >1 samples with non-zero read count ####  >>> NEW (in the original script this was being done individually for each stage)
countData <- countData[rowSums(countData > 0) >= 2 , ]
nrow(countData)



#Make separate data sets for each developmental stage ####
All_body_countdata <- countData[ ,-c(66:71)]
All_body_coldata <- as.data.frame(colData[c(1:65), ])
head(All_body_countdata)

Instar5body_countdata <- subset(countData, select = c("P11917_102_S50", "P11917_103_S51", "P11917_105_S53", "P11917_106_S54", 
                                                      "P11917_108_S56", "P11917_109_S57", "P11917_111_S59", "P11917_112_S60", "P11917_114_S62", 
                                                      "P11917_115_S63", "P11917_117_S65", "P11917_118_S66", "P11917_120_S68", "P11917_121_S69", 
                                                      "P11917_123_S70", "P11917_124_S71", "P11917_126_S73", "P11917_127_S74", "P11917_129_S76", 
                                                      "P11917_130_S77", "P11917_132_S79", "P11917_133_S80", "P11917_135_S82", "P11917_136_S83", 
                                                      "P11917_138_S85", "P11917_139_S86", "P11917_141_S88", "P11917_142_S89", "P11917_144_S91", 
                                                      "P11917_145_S92", "P11917_147_S94", "P11917_148_S95", "P11917_150_S97", "P11917_151_S98", 
                                                      "P11917_153_S100", "P11917_154_S101", "P11917_156_S103", "P11917_157_S104", "P11917_159_S106", 
                                                      "P11917_160_S107", "P11917_162_S109", "P11917_163_S110", "P11917_165_S112", "P11917_166_S113"))
Instar5body_coldata <- as.data.frame(colData[-c(1, 4, 7, 10, 13, 16, 19, 24, 27, 30, 33, 36, 39, 42, 45, 48, 51, 54, 57, 60, 63, 66:71), ])

Instar3_countdata <- subset(countData, select = c("P11917_101_S49", "P11917_104_S52", "P11917_107_S55", 
                                                 "P11917_110_S58", "P11917_113_S61", "P11917_116_S64", "P11917_119_S67", "P11917_125_S72", 
                                                 "P11917_128_S75", "P11917_131_S78", "P11917_134_S81", "P11917_137_S84", "P11917_140_S87", 
                                                 "P11917_143_S90", "P11917_146_S93", "P11917_149_S96", "P11917_152_S99", "P11917_155_S102", 
                                                 "P11917_158_S105", "P11917_161_S108", "P11917_164_S111"))
Instar3_coldata <- as.data.frame(colData[c(1, 4, 7, 10, 13, 16, 19, 24, 27, 30, 33, 36, 39, 42, 45, 48, 51, 54, 57, 60, 63), ])

Instar5M_countdata <- subset(countData, select = c("P11917_102_S50", "P11917_105_S53", "P11917_108_S56", "P11917_111_S59", "P11917_114_S62", 
                                                   "P11917_117_S65", "P11917_120_S68", "P11917_123_S70", "P11917_126_S73", "P11917_129_S76", 
                                                   "P11917_132_S79", "P11917_135_S82", "P11917_138_S85", "P11917_141_S88", "P11917_144_S91", 
                                                   "P11917_147_S94", "P11917_150_S97", "P11917_153_S100", "P11917_156_S103", "P11917_159_S106", 
                                                   "P11917_162_S109", "P11917_165_S112"))

Instar5M_coldata <- as.data.frame(colData[c(2, 5, 8, 11, 14, 17, 20, 22, 25, 28, 31, 34, 37, 40, 43, 46, 49, 52, 55, 58, 61, 64), ])

Instar5F_countdata <- subset(countData, select = c("P11917_103_S51", "P11917_106_S54", "P11917_109_S57", "P11917_112_S60", "P11917_115_S63", 
                                                   "P11917_118_S66", "P11917_121_S69", "P11917_124_S71", "P11917_127_S74", "P11917_130_S77", 
                                                   "P11917_133_S80", "P11917_136_S83", "P11917_139_S86", "P11917_142_S89", "P11917_145_S92", 
                                                   "P11917_148_S95", "P11917_151_S98", "P11917_154_S101", "P11917_157_S104", "P11917_160_S107", 
                                                   "P11917_163_S110", "P11917_166_S113"))
Instar5F_coldata <- as.data.frame(colData[c(3, 6, 9, 12, 15, 18, 21, 23, 26, 29, 32, 35, 38, 41, 44, 47, 50, 53, 56, 59, 62, 65), ])

Instar3_coldata <- Instar3_coldata[ , 1:6]
Instar5M_coldata <- Instar5M_coldata[ , 1:6]
Instar5F_coldata <- Instar5F_coldata[ , 1:6]



```


#Filtering the data by removing genes with 0 counts in each dataset (first filtering removing all genes with 0 in the whole dataset done before partitioning the data). 
1 count is added to the whole data set after filtering 0 counts genes. 
Chech rows and colomns. 

```{r filtering}
#Pre-filtering II: for each developmental stage, remove genes with zero counts in all samples ####  
Instar3_countdata <- Instar3_countdata[rowSums(Instar3_countdata > 0) >= 1 , ]
nrow(Instar3_countdata)
Instar5M_countdata <- Instar5M_countdata[rowSums(Instar5M_countdata > 0) >= 1 , ]
nrow(Instar5M_countdata)
Instar5F_countdata <- Instar5F_countdata[rowSums(Instar5F_countdata > 0) >= 1 , ]
nrow(Instar5F_countdata)
All_body_countdata <- All_body_countdata[rowSums(All_body_countdata > 0) >= 1 , ]
nrow(All_body_countdata)
Instar5body_countdata <- Instar5body_countdata[rowSums(Instar5body_countdata > 0) >=1 , ]
nrow(Instar5body_countdata)


## Add 1 count to every gene/sample (multiple zero-count samples take DESEQ off track)
Instar3_countdata <- Instar3_countdata + 1
head(Instar3_countdata)
Instar5M_countdata <- Instar5M_countdata + 1
head(Instar5M_countdata)
Instar5F_countdata <- Instar5F_countdata + 1
head(Instar5F_countdata)
All_body_countdata <- All_body_countdata + 1
head(All_body_countdata)
Instar5body_countdata <- Instar5body_countdata + 1
head(Instar5body_countdata)

#Check all sample IDs in colData are also in CountData and match their orders ####
all(rownames(colData) %in% colnames(countData))
countData <- countData[, rownames(colData)]
all(rownames(colData) == colnames(countData))

all(rownames(All_body_coldata) %in% colnames(All_body_countdata))
All_body_countdata <- All_body_countdata[, rownames(All_body_coldata)]
all(rownames(All_body_coldata) == colnames(All_body_countdata))

all(rownames(Instar3_coldata) %in% colnames(Instar3_countdata))
Instar3_countdata <- Instar3_countdata[, rownames(Instar3_coldata)]
all(rownames(Instar3_coldata) == colnames(Instar3_countdata))

all(rownames(Instar5M_coldata) %in% colnames(Instar5M_countdata))
Instar5M_countdata <- Instar5M_countdata[, rownames(Instar5M_coldata)]
all(rownames(Instar5M_coldata) == colnames(Instar5M_countdata))

all(rownames(Instar5F_coldata) %in% colnames(Instar5F_countdata))
Instar5F_countdata <- Instar5F_countdata[, rownames(Instar5F_coldata)]
all(rownames(Instar5F_coldata) == colnames(Instar5F_countdata))

all(rownames(Instar5body_coldata) %in% colnames(Instar5body_countdata))
Instar5body_countdata <- Instar5body_countdata[, rownames(Instar5body_coldata)]
all(rownames(Instar5body_coldata) == colnames(Instar5body_countdata))

```

#Creating DE datasets
design is the set up of the model Variables + ev interaction term

```{r DE_datasets}
#Create a DESeqDataSet from count matrix and labels (no confounding effects removed at this stage) ####
ddsAll_Pop_Host <- DESeqDataSetFromMatrix(countData = countData,
                                    colData = colData, design = ~Population+Hostplant)
ddsAll_Pop_Host

ddsAll_Body <- DESeqDataSetFromMatrix(countData = All_body_countdata,
                                          colData = All_body_coldata, design = ~Population+Hostplant+Stage)
ddsAll_Body

ddsAll_Body_Host <- DESeqDataSetFromMatrix(countData = All_body_countdata,
                                          colData = All_body_coldata, design = ~Population+Hostplant)
ddsAll_Body_Host

ddsAll_Body_interaction <- DESeqDataSetFromMatrix(countData = All_body_countdata,
                                          colData = All_body_coldata, design = ~Population+Hostplant+Stage+Population:Hostplant)



ddsInstar5_Body <- DESeqDataSetFromMatrix(countData = Instar5body_countdata,
                                      colData = Instar5body_coldata, design = ~Hostplant+Population)
ddsInstar5_Body

ddsInstar5_Body2 <- DESeqDataSetFromMatrix(countData = Instar5body_countdata,
                                      colData = Instar5body_coldata, design = ~Hostplant+Population+Sex)
ddsInstar5_Body2

ddsInstar3_Pop_Host <- DESeqDataSetFromMatrix(countData = Instar3_countdata,
                                    colData = Instar3_coldata, design = ~Population+Hostplant)
ddsInstar3_Pop_Host

ddsInstar5M_Pop_Host <- DESeqDataSetFromMatrix(countData = Instar5M_countdata,
                                  colData = Instar5M_coldata, design = ~Population+Hostplant)
ddsInstar5M_Pop_Host

ddsInstar5F_Pop_Host <- DESeqDataSetFromMatrix(countData = Instar5F_countdata,
                                   colData = Instar5F_coldata, design = ~Population+Hostplant)
ddsInstar5F_Pop_Host


```

#Filtering again

```{r filtering2}
#Pre-filtering IIIs: Only keep rows with baseMean >2 (usually this woulf be baseMean >1 but we added one count to every gene/sample)

ddsAll_Pop_Host <- ddsAll_Pop_Host[rowMeans(counts(ddsAll_Pop_Host)) > 2, ]

ddsAll_Body <- ddsAll_Body[rowMeans(counts(ddsAll_Body)) > 2, ]
#ddsAll_Body
ddsAll_Body_Host <- ddsAll_Body_Host[rowMeans(counts(ddsAll_Body_Host)) > 2, ]

ddsInstar5_Body <- ddsInstar5_Body[rowMeans(counts(ddsInstar5_Body)) > 2, ]
#ddsInstar5_Body

ddsInstar3_Pop_Host <- ddsInstar3_Pop_Host[rowMeans(counts(ddsInstar3_Pop_Host)) > 2, ]
#ddsInstar3_Pop_Host
ddsInstar5M_Pop_Host <- ddsInstar5M_Pop_Host[rowMeans(counts(ddsInstar5M_Pop_Host)) > 2, ]
#ddsInstar5M_Pop_Host
ddsInstar5F_Pop_Host <- ddsInstar5F_Pop_Host[rowMeans(counts(ddsInstar5F_Pop_Host)) > 2, ]
#ddsInstar5F_Pop_Host

```

#Running the actual DE analysis

Extract the results by using contrast for each comparison

```{r DEanalysis}
# Without batch effect correction
#ddsAll_Pop_Host <- DESeq(ddsAll_Pop_Host)
#res_All_Pop_Host <- results(ddsAll_Pop_Host, alpha =.05)
#res_All_Pop_Host
#summary(res_All_Pop_Host)

#res_All_Pop_Host_contr <- results(ddsAll_Pop_Host, alpha =.05, contrast = c("Population", "span", "swe") )
#res_All_Pop_Host_contr
#summary(res_All_Pop_Host_contr)

ddsAll_Body <- DESeq(ddsAll_Body)
res_All_Body <- results(ddsAll_Body, alpha =.05)
res_All_Body
summary(res_All_Body)

ddsAll_Body_Host <- DESeq(ddsAll_Body_Host)
res_All_Body_Host <- results(ddsAll_Body_Host, alpha =.05)
res_All_Body_Host
summary(res_All_Body_Host)



res_All_Body_contr <- results(ddsAll_Body, alpha =.05, contrast = c("Population", "span", "swe") )
#res_All_Body_contr
summary(res_All_Body_contr)

res_All_Body_contr_plant <- results(ddsAll_Body, alpha =.05, contrast = c("Hostplant", "Lot", "Dor") )
#res_All_Body_contr_plant
summary(res_All_Body_contr_plant)

ddsAll_Body_interaction <- DESeq(ddsAll_Body_interaction)
res_All_Body_interaction <- results(ddsAll_Body_interaction,alpha = .05)
res_All_Body_interaction
summary(res_All_Body_interaction)

res_All_Body_interaction_hostplant <- results(ddsAll_Body_interaction,alpha = .05, contrast = c("Hostplant", "Lot", "Dor"))
res_All_Body_interaction_hostplant
summary(res_All_Body_interaction_hostplant)



#not working?
res_All_Body_interaction_popLot <- results(ddsAll_Body_interaction,alpha = .05, contrast = c("Population", "span", "swe"))
res_All_Body_interaction_popLot
summary(res_All_Body_interaction_popLot)



ddsInstar5_Body <- DESeq(ddsInstar5_Body)
res_Instar5_Body <- results(ddsInstar5_Body, alpha =.05)
#res_Instar5_Body
summary(res_Instar5_Body)

res_Instar5_Body_contr_plant <- results(ddsInstar5_Body, alpha =.05, contrast = c("Hostplant", "Lot", "Dor") )
summary(res_Instar5_Body_contr_plant)

ddsInstar5_Body2 <- DESeq(ddsInstar5_Body2)
res_Instar5_Body2 <- results(ddsInstar5_Body2, alpha =.05)
#res_Instar5_Body2
summary(res_Instar5_Body2)

res_Instar5_Body2_hostplant <- results(ddsInstar5_Body2, alpha =.05, contrast= c("Hostplant", "Lot", "Dor"))
#res_Instar5_Body2_hostplant
summary(res_Instar5_Body2_hostplant)

res_Instar5_Body2_pop <- results(ddsInstar5_Body2, alpha =.05, contrast= c("Population", "span", "swe"))
#res_Instar5_Body2_pop
summary(res_Instar5_Body2_pop)



ddsInstar3_Pop_Host <- DESeq(ddsInstar3_Pop_Host)
res_Instar3_Pop_Host <- results(ddsInstar3_Pop_Host, alpha=.05)
summary(res_Instar3_Pop_Host)
res_Instar3_contr1_Pop <- results(ddsInstar3_Pop_Host, alpha=.05, contrast = c("Population", "span", "swe"))
summary(res_Instar3_contr1_Pop)

ddsInstar5M_Pop_Host <- DESeq(ddsInstar5M_Pop_Host)
res_Instar5M_Pop_Host <- results(ddsInstar5M_Pop_Host, alpha=.05)
summary(res_Instar5M_Pop_Host)

res_Instar5M_contr1_Pop <- results(ddsInstar5M_Pop_Host, alpha=.05, contrast = c("Population", "span", "swe"))
summary(res_Instar5M_contr1_Pop)

ddsInstar5F_Pop_Host <- DESeq(ddsInstar5F_Pop_Host)
res_Instar5F_Pop_Host <- results(ddsInstar5F_Pop_Host, alpha=.05)
summary(res_Instar5F_Pop_Host)

res_Instar5F_contr1_Pop <- results(ddsInstar5F_Pop_Host, alpha=.05, contrast = c("Population", "span", "swe"))
summary(res_Instar5F_contr1_Pop)

```


```{r filter III}
#filter result only keep basemean over 10 
#
#Filtered_P0.05_base_Mean_10_Instar <- subset(Filtered_P0.05_Instar3_Pop_Host, baseMean>10
Filtered_BM10_res_all_body <- subset(res_All_Body, baseMean>10)
Filtered_BM10_res_all_body_host <- subset(res_All_Body_Host, baseMean>10)



```



#Visualisation of the data

```{r visualisation}
#plotting the results
#MA plot
par(mfrow=c(2,2))
plotMA(res_All_Body_contr_plant, main= "All, Lot vs Dor")
plotMA(res_Instar3_Pop_Host, main="Instar3, Lot vs Dor")
plotMA(res_Instar5M_Pop_Host, main="Instar5M, Lot vs Dor")
plotMA(res_Instar5F_Pop_Host, main="Instar5F, Lot vs Dor")
plotMA(res_All_Body_contr, main="All, span vs swe")
plotMA(res_Instar3_contr1_Pop, main="Instar3, span vs swe")
plotMA(res_Instar5M_contr1_Pop, main="Instar5M, span vs swe")
plotMA(res_Instar5F_contr1_Pop, main="Instar5F, span vs swe")

plotCounts(ddsAll_Body, gene=which.min(res_All_Body$padj), intgroup="Hostplant", xlab = "All")
plotCounts(ddsInstar3_Pop_Host, gene=which.min(res_Instar3_Pop_Host$padj), intgroup="Hostplant", xlab = "Instar 3")
plotCounts(ddsInstar5M_Pop_Host, gene=which.min(res_Instar5M_Pop_Host$padj), intgroup="Hostplant", xlab = "Instar 5 Male")
plotCounts(ddsInstar5F_Pop_Host, gene=which.min(res_Instar5F_Pop_Host$padj), intgroup="Hostplant", xlab = "Instar 5 Female")

#heatmap
#install.packages("pheatmap")
library("pheatmap")

#try sorting in pheatmap command, it works!
All_Body.ntd <- normTransform(ddsAll_Body)
select_All_Body <- order(rowMeans(counts(ddsAll_Body,normalized=TRUE)),
                             decreasing=TRUE)

All_Body.df <- as.data.frame(colData(ddsAll_Body)[,c("Population", "Hostplant")])
pheatmap(assay(All_Body.ntd)[select_All_Body,order(All_Body.ntd$Hostplant)], cluster_rows=FALSE, show_rownames=FALSE,
         cluster_cols=FALSE, annotation_col=All_Body.df)

#cluster columns
All_Body.df <- as.data.frame(colData(ddsAll_Body)[,c("Population", "Hostplant", "Stage")])
pheatmap(assay(All_Body.ntd)[select_All_Body,order(All_Body.ntd$Hostplant)], cluster_rows=FALSE, show_rownames=FALSE,
         cluster_cols=TRUE, annotation_col=All_Body.df)



#adds a clustertree to columns
# pheatmap(assay(All_Body.ntd)[select_All_Body,], cluster_rows=FALSE, show_rownames=FALSE,
#          cluster_cols=TRUE, annotation_col=All_Body.df)
# 
# #adds a clustertree to rows
# pheatmap(assay(All_Body.ntd)[select_All_Body,], cluster_rows=TRUE, show_rownames=FALSE,
#          cluster_cols=FALSE, annotation_col=All_Body.df)
# 
# #not working
# All_Body2.df <- as.data.frame(colData(ddsAll_Body)[,"Stage"])
# pheatmap(assay(All_Body.ntd)[select_All_Body,], cluster_rows=FALSE, show_rownames=FALSE,
#          cluster_cols=FALSE, annotation_col=All_Body2.df)
# 
# # changes order of pop and hostplant but not clustering
# All_Body2.df <- as.data.frame(colData(ddsAll_Body)[,c("Population", "Hostplant")])
# pheatmap(assay(All_Body.ntd)[select_All_Body,], cluster_rows=FALSE, show_rownames=FALSE,
#          cluster_cols=FALSE, annotation_col=All_Body2.df)


#Instar5
Instar5_Body.ntd <- normTransform(ddsInstar5_Body)
select_Instar5_Body <- order(rowMeans(counts(ddsInstar5_Body,normalized=TRUE)),
                             decreasing=TRUE)
Instar5_Body.df <- as.data.frame(colData(ddsInstar5_Body)[,c("Population", "Hostplant")])
pheatmap(assay(Instar5_Body.ntd)[select_Instar5_Body,order(Instar5_Body.ntd$Hostplant)], cluster_rows=FALSE, show_rownames=FALSE,
         cluster_cols=FALSE, annotation_col=Instar5_Body.df)

#cluster columns
Instar5_Body.ntd <- normTransform(ddsInstar5_Body)
select_Instar5_Body <- order(rowMeans(counts(ddsInstar5_Body,normalized=TRUE)),
                             decreasing=TRUE)
Instar5_Body.df <- as.data.frame(colData(ddsInstar5_Body)[,c("Population", "Hostplant")])
pheatmap(assay(Instar5_Body.ntd)[select_Instar5_Body,order(Instar5_Body.ntd$Hostplant)], cluster_rows=FALSE, show_rownames=FALSE,
         cluster_cols=TRUE, annotation_col=Instar5_Body.df)

#Instar3
Instar3.ntd <- normTransform(ddsInstar3_Pop_Host)
select_Instar3 <- order(rowMeans(counts(ddsInstar3_Pop_Host,normalized=TRUE)),
                             decreasing=TRUE)
Instar3.df <- as.data.frame(colData(ddsInstar3_Pop_Host)[,c("Population", "Hostplant")])
pheatmap(assay(Instar3.ntd)[select_Instar3,order(Instar3.ntd$Hostplant)], cluster_rows=FALSE, show_rownames=FALSE,
         cluster_cols=FALSE, annotation_col=Instar3.df)

# #cluster columns
# #Instar3
Instar3.ntd <- normTransform(ddsInstar3_Pop_Host)
select_Instar3 <- order(rowMeans(counts(ddsInstar3_Pop_Host,normalized=TRUE)),
                             decreasing=TRUE)
Instar3.df <- as.data.frame(colData(ddsInstar3_Pop_Host)[,c("Population", "Hostplant")])
pheatmap(assay(Instar3.ntd)[select_Instar3,order(Instar3.ntd$Hostplant)], cluster_rows=FALSE, show_rownames=FALSE,
         cluster_cols=TRUE, annotation_col=Instar3.df)
```





```{r pca}

#PCA
library(ggplot2)
All_Body.vsd <- varianceStabilizingTransformation(ddsAll_Body, blind = T)
plotPCA(All_Body.vsd, intgroup=c("Hostplant","Population")) + theme_minimal() + ggtitle("PCA All")


# All_Body_Host.vsd <- varianceStabilizingTransformation(ddsAll_Body_Host, blind = T)
# plotPCA(All_Body_Host.vsd, intgroup=c("Hostplant","Population")) + theme_minimal() + ggtitle("PCA All_Host")

plotPCA(All_Body.vsd, intgroup=c("Hostplant", "Population", "Stage")) + 
     geom_point(aes(colour=Hostplant, fill=Population, shape = Stage), size = 4) +
     theme_minimal() + 
     scale_fill_manual(values = c("transparent", "white")) +
 scale_shape_manual(values = c(1,6))

#use only the data from plotPCA and then plot with ggplot
All_body_pca<-plotPCA(All_Body.vsd, intgroup=c("Hostplant","Population", "Stage"), returnData = T)
head(All_body_pca)
write.table(All_body_pca, file = "all_body_pca.txt")


All_body_pca1 <- read.table(file = "all_body_pca.txt")


All_body_pca <- read.table("all_body_pca.txt")
#size is added just to get the legend right 
ggplot(All_body_pca, aes(x=PC1, 
                         y=PC2,
                         colour = Hostplant,
                         fill=interaction(Population,Hostplant),
                         shape = Stage, 
                         size = Population)) +
  geom_point(stroke = 1) +
  scale_fill_manual(values = c("orangered4", "transparent", "orange2", "transparent"),
                    guide = "none") +
  scale_colour_manual(values = c("orangered4", "orange2"),
                      labels = c("L. dorycnium", "L. corniculatus"),
                      guide =
                      guide_legend(label.theme = element_text(face = "italic"),
                                   order = 1,
                                   override.aes = list(size = 4))) +
  scale_shape_manual(values = c(21,25),
                     labels = c("III", "V"),
                     guide =
                       guide_legend(override.aes = list(size = 4),
                                    order = 2)) +
  scale_size_manual(name = "Population",
                    values = c(4, 4.01), 
                    labels = c("Catalonia", "Sweden"),
                    guide = guide_legend(override.aes = list(shape = c(16,1)),
                                         order = 2)) +
  labs(x = "PC1: 22% variance", y = "PC2: 18% variance") +
  theme(panel.border = element_blank(), 
          axis.line = element_line(color = 'black'),
          axis.text.x = element_text(size = 12),
        axis.title.x = element_text(size = 12),
          axis.text.y = element_text(size = 12),
        axis.title.y = element_text(size = 12),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), 
        legend.key = element_blank(),
          legend.text = element_text(size = 12))
  
        #legend.text = element_text(text() angle = 0, size = 12, face = "italic)) #green version
ggplot(All_body_pca, aes(x=PC1, 
                         y=PC2,
                         colour = Hostplant,
                         fill=interaction(Population,Hostplant),
                         shape = Stage, 
                         size = Population)) +
  geom_point(stroke = 1) +
  scale_fill_manual(values = c("chartreuse4", "transparent", "chartreuse2", "transparent"),
                    guide = "none") +
  scale_colour_manual(values = c("chartreuse4", "chartreuse2"),
                      labels = c("L. dorycnium", "L. corniculatus"),
                      guide =
                      guide_legend(label.theme = element_text(face = "italic"),
                                   order = 1,
                                   override.aes = list(size = 4))) +
  scale_shape_manual(values = c(21,25),
                     labels = c("III", "V"),
                     guide =
                       guide_legend(override.aes = list(size = 4),
                                    order = 2)) +
  scale_size_manual(name = "Population",
                    values = c(4, 4.01), 
                    labels = c("Catalonia", "Sweden"),
                    guide = guide_legend(override.aes = list(shape = c(16,1)),
                                         order = 2)) +
  labs(title = "All", x = "PC1: 22% variance", y = "PC2: 18% variance") +
  xlim(-80, 90) +
  ylim(-50, 75) +
  theme(panel.border = element_blank(), 
        axis.line = element_line(color = 'black'),
        axis.text.x = element_text(size = 12),
        axis.title.x = element_text(size = 16),
        axis.text.y = element_text(size = 12),
        axis.title.y = element_text(size = 16),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        legend.key = element_blank(),
        legend.text = element_text(size = 12), 
        legend.title = element_text(size = 16), 
        title = element_text(size = 20))
  

  
























All_Body_interaction.vsd <- varianceStabilizingTransformation(ddsAll_Body_interaction, blind = T)
plotPCA(All_Body_interaction.vsd, intgroup=c("Hostplant","Population")) + theme_minimal() + ggtitle("PCA All Interaction")

par(mfrow=c(2,2))
plotPCA(All_Body.vsd, intgroup=c("Hostplant","Population", "Stage")) + theme_minimal()
plotPCA(All_Body.vsd, intgroup=c("Stage")) + theme_minimal() + ggtitle("PCA All Instar 3 vs 5")
plotPCA(All_Body.vsd, intgroup=c("Population")) + theme_minimal() + ggtitle("PCA All Population")
plotPCA(All_Body.vsd, intgroup=c("Hostplant")) + theme_minimal() + ggtitle("PCA All Hostplant")

Instar5_Body.vsd <- varianceStabilizingTransformation(ddsInstar5_Body, blind = T)
pcaInstar5_body <- plotPCA(Instar5_Body.vsd, intgroup=c("Population", "Hostplant")) + ggtitle("PCA Instar 5")
plotPCA(Instar5_Body.vsd, intgroup=c("Hostplant")) + theme_minimal() + ggtitle("PCA Instar 5 Hostplant")
plotPCA(Instar5_Body.vsd, intgroup=c("Population")) + theme_minimal() + ggtitle("PCA Instar 5 Population")
plotPCA(Instar5_Body.vsd, intgroup=c("Sex")) + theme_minimal() + ggtitle("PCA Instar 5 Sex")
plotPCA(Instar5_Body.vsd, intgroup=c("Family")) + theme_minimal() + ggtitle("PCA Instar 5 Family")

InstarV_pca<-plotPCA(Instar5_Body.vsd, intgroup=c("Hostplant","Population"), returnData = T)
head(InstarV_pca)
write.table(InstarV_pca, file = "InstarV_pca.txt")

InstarV_pca <- read.table("InstarV_pca.txt")

ggplot(InstarV_pca, aes(x=PC1, 
                         y=PC2,
                         colour = Hostplant,
                         fill=interaction(Population,Hostplant),
                         size = Population)) +
  geom_point(shape = 25, stroke = 1) +
  scale_fill_manual(values = c("chartreuse4", "transparent", "chartreuse2", "transparent"),
                    guide = "none") +
  scale_colour_manual(values = c("chartreuse4", "chartreuse2"),
                      labels = c("L. dorycnium", "L. corniculatus"),
                      guide =
                      guide_legend(label.theme = element_text(face = "italic"),
                                   order = 1,
                                   override.aes = list(size = 4, shape = 16))) +
  # #scale_shape_manual(values = c(21,25),
  #                    labels = c("III", "V"),
  #                    guide =
  #                      guide_legend(override.aes = list(size = 4),
  #                                   order = 2)) +
  scale_size_manual(name = "Population",
                    values = c(4, 4.01), 
                    labels = c("Catalonia", "Sweden"),
                    guide = guide_legend(override.aes = list(shape = c(16,1)),
                                         order = 2)) +
  labs(title = "InstarV", x = "PC1: 27% variance", y = "PC2: 17% variance") +
  xlim(-80, 90) +
  ylim(-50, 75) +
  theme(panel.border = element_blank(), 
          axis.line = element_line(color = 'black'),
          axis.text.x = element_text(size = 12),
        axis.title.x = element_text(size = 16),
          axis.text.y = element_text(size = 12),
        axis.title.y = element_text(size = 16),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), 
        legend.key = element_blank(),
          legend.text = element_text(size = 12), 
        legend.title = element_text(size = 16),
        title = element_text(size = 20))
  

Instar3.vsd <- varianceStabilizingTransformation(ddsInstar3_Pop_Host, blind = T)
plotPCA(Instar3.vsd, intgroup=c("Population", "Hostplant")) + theme_minimal() + ggtitle("PCA Instar 3")
plotPCA(Instar3.vsd, intgroup=c("Hostplant")) + theme_minimal() + ggtitle("PCA Instar 3 Hostplant")
plotPCA(Instar3.vsd, intgroup=c("Population")) + theme_minimal() + ggtitle("PCA Instar 3 Population")
plotPCA(Instar3.vsd, intgroup=c("Family")) + theme_minimal() + ggtitle("PCA Instar 3 Family")

InstarIII_pca<-plotPCA(Instar3.vsd, intgroup=c("Hostplant","Population"), returnData = T)
head(InstarIII_pca)
write.table(InstarIII_pca, file = "InstarIII_pca.txt")

InstarIII_pca <- read.table("InstarIII_pca.txt")


ggplot(InstarIII_pca, aes(x=PC1, 
                         y=PC2,
                         colour = Hostplant,
                         fill=interaction(Population,Hostplant),
                         size = Population)) +
  geom_point(shape = 21, stroke = 1) +
  scale_fill_manual(values = c("chartreuse4", "transparent", "chartreuse2", "transparent"),
                    guide = "none") +
  scale_colour_manual(values = c("chartreuse4", "chartreuse2"),
                      labels = c("L. dorycnium", "L. corniculatus"),
                      guide =
                      guide_legend(label.theme = element_text(face = "italic"),
                                   order = 1,
                                   override.aes = list(size = 4, shape = 16))) +
  # #scale_shape_manual(values = c(21,25),
  #                    labels = c("III", "V"),
  #                    guide =
  #                      guide_legend(override.aes = list(size = 4),
  #                                   order = 2)) +
  scale_size_manual(name = "Population",
                    values = c(4, 4.01), 
                    labels = c("Catalonia", "Sweden"),
                    guide = guide_legend(override.aes = list(shape = c(16,1)),
                                         order = 2)) +
  labs(title = "InstarIII", x = "PC1: 42% variance", y = "PC2: 13% variance") +
  xlim(-80, 90) +
  ylim(-50, 75) +
  theme(panel.border = element_blank(), 
          axis.line = element_line(color = 'black'),
          axis.text.x = element_text(size = 12),
        axis.title.x = element_text(size = 16),
          axis.text.y = element_text(size = 12),
        axis.title.y = element_text(size = 16),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(), 
        legend.key = element_blank(),
          legend.text = element_text(size = 12), 
        legend.title = element_text(size = 16),
        title = element_text(size = 20))


```
